---
title: "EDS 211: Team Science"
description: "EDS 211: Team Science, Collaborative Analysis and Project Management"
site: distill::distill_website
bibliography: ["eds211-team.bib"]
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F)
```

> If you want to go fast, go alone. If you want to go far, go together. -- African proverb

# Schedule

```{r}
source(here::here("functions.R"))
get_sched() %>% 
  dt_sched()
```

# Overview

Welcome to the **Team Science** course for the Masters in Environmental Data Science! In this class you'll learn how to work together on code while managing a project. You'll apply these techniques to the group project from the course [Remote Sensing (EDS 220)](https://github.com/samanthastevenson/EDS220_Fall2021) and they should help you manage your Capstone group project starting winter quarter.
The techniques will help guide three basic questions:

1.  What should I work on next?
2.  How long will it take?
3.  How do we adapt to change?

These basic questions are however deceptively difficult to answer, especially as the size of the team and scope of the project increases.

**What should I work on next?** You'll learn 'agile' software development techniques to divide work into tasks, which can be assigned to various team members and passed between as work progresses. These tasks can be arranged as cards into columns on a 'Kanban' board: horizontally according to status (ToDo, In Progress, Done) or time (next week, next month, ...); and vertically based on priority (from most to least important, top to bottom). This overview of who is to do what next will be approached with different tools embracing increasing levels of complexity: a Trello board for basic projects without code, a Github Project to integrate Github Issues associated with code level git commit messages, and ZenHub for adding dependencies between issues and clustering into Epics for use with Roadmaps.

**How long will it take?** You'll take on the role of an environmental data science consultancy responding to a request for proposal. In the proposal you'll outline milestones with deliverables and associated dollar amounts based on time estimates and hourly rates of team members. You'll track time spent on each task and report out at the end with a presentation on time spent over and under expectations, which will help update your understanding for future estimates. You'll use Clockify to track time and Rmarkdown to generate reports.

**How do we adapt to change?** Changes are inevitable with technical problem solving and client requirements. As Mark Zuckerberg famously said "move fast and break things." Iterating quickly within the team and getting feedback from all stakeholders are an important part of the development process and the basis for 'agile' project management. You'll compare 'traditional' project management with scoping tasks at the outset of the project with a Gantt chart as a 'waterfall' diagram showing various phases of the project (eg data collection, analysis and communication).
As you proceed with the project you'll keep track of tasks that expand and shift with some evaluation after the fact through a burndown report. Other 'agile' techniques and roles will be reviewed to adapt to change.

# Grading

Grades will be based on the following percentages:

- Attendance: 10%
- Labs: 60%
- Presentation: 30%

# Readings

```{r, eval=F}
library(citr)
md_cite("Bennett", bib_file="eds211-team.bib")
```

<!-- TODO: move to UCSB only Google Drive -->
- Week 1: @bennettCollaborationTeamScience2012 ([pdf](https://drive.google.com/file/d/1LJWfJTo4EWcd6Vxyfg7R6GTES-orY-3B/view?usp=sharing))
- Week 2: @lowndesOurPathBetter2017 ([pdf](https://drive.google.com/file/d/16eswlI2SFQpITD1Oqbi7NPDhTm3Xope4/view?usp=sharing))
- Week 3: @perez-riverolTenSimpleRules2016 ([pdf](https://drive.google.com/file/d/1gt0jrZ3Q6HCqSORb7_L24Twj7WiIlCwO/view?usp=sharing))
- Week 4: @larsonBestPracticesAccelerating2019 ([pdf](https://drive.google.com/file/d/1ZnAK5c-96B8G_dZkNA3SD6qppEuWz5iA/view?usp=sharing))


